## WorldFootBallR Package

```{r}
# install.packages("worldfootballR")
library(worldfootballR)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(plotly)
library(dplyr)
library(ggrepel)
library(gtable)
```

## User Case 1: Analyzing Forward Players' Performance Over Time of FC Barcelona (from fbref.com) 

```{r}
# fc_barcelona_url <- "https://fbref.com/en/squads/206d90db/2022-2023/Barcelona-Stats"
# match_logs <- list()
# player_urls <- fb_player_urls(team_url = fc_barcelona_url)
# for (player_url in player_urls) {
#   logs <- fb_player_match_logs(player_url = player_url, season_end_year = 2023, stat_type = "summary")
#   match_logs <- append(match_logs, list(logs))
#   Sys.sleep(60)
# }
# all_match_logs <- bind_rows(match_logs)
# write.csv(all_match_logs, "match_logs.csv", row.names = FALSE)
```

```{r}
existing_logs <- read.csv("match_logs.csv")
head(existing_logs)
```

### Filtering only players who play in forward positions

```{r}
colnames(existing_logs) <- trimws(colnames(existing_logs))
forwards <- existing_logs |>
  filter(existing_logs$Pos == "FW")
```

### Selecting Key Performance Attributes for Forwards to Analyze

```{r}
forwards$Date <- as.Date(forwards$Date, format = "%m/%d/%y")  
forwards$Date[is.na(forwards$Date)] <- as.Date(forwards$Date[is.na(forwards$Date)], 
                                               format = "%Y-%m-%d")  

forwards_performance <- forwards |>
  select(Date, Gls_Performance, Ast_Performance, SoT_Performance, xG_Expected)
head(forwards_performance)
```

### Analyzing the different performance metrics for various forwards' in FC Barcelona

```{r}

# Plot with legend for goals performance
get_legend <- function(my_plot) {
  tmp <- ggplotGrob(my_plot)
  leg <- gtable_filter(tmp, "guide-box")
  return(leg)
}

# Plotting Goals 
plot_with_legend <- ggplot(forwards, aes(x = Date, 
                                         y = Gls_Performance, color = Player)) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  labs(title = "Goals for Forward Players", y = "Goals", x = "Date") +
  theme_minimal() +
  theme(
    legend.position = "bottom", 
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
    plot.title = element_text(size = 10),
    axis.title = element_text(size = 8)
  ) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 month")

common_legend <- get_legend(plot_with_legend)
plot_goals <- plot_with_legend + theme(legend.position = "none")

```

```{r}
# Plot for assists
plot_assists <- ggplot(forwards, aes(x = Date, y = Ast_Performance, color = Player)) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  labs(title = "Assists for Forward Players", y = "Assists", x = "Date") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 month")
```

```{r}
# Plot for expected goals (xG)
plot_expected_goals <- ggplot(forwards, aes(x = Date, y = xG_Expected, color = Player)) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  labs(title = "Expected Goals (xG) for Forward Players", y = "Expected Goals (xG)", x = "Date") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 month")
```

```{r}
# Plot for shots on target
plot_shots_on_target <- ggplot(forwards, aes(x = Date, y = SoT_Performance, color = Player)) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  labs(title = "Shots on Target for Forward Players", y = "Shots on Target", x = "Date") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 8)) +
  scale_x_date(date_labels = "%Y-%m", date_breaks = "1 month")
```

```{r, fig.width = 11, fig.height = 10}
# Arrange plots with a common legend at the bottom
grid.arrange(
  arrangeGrob(plot_goals, plot_assists, plot_shots_on_target,
              plot_expected_goals, 
              ncol = 2,
              top = "Performance Metrics Over Time for Forward Players (2022-2023)"),
  common_legend,
  nrow = 2,
  heights = c(10, 1) 
)
```

## Analyzing Performance of a Single Player (Robert Lewandowski) for 2021-2023

```{r}
# Getting Robert Lewandowski's match logs from 2021-2023
robert_lewandowski <- forwards |>
  filter(Player == "Robert Lewandowski")
robert_url <- "https://fbref.com/en/players/8d78e732/Robert-Lewandowski"
robert_stats_2021 <- fb_player_match_logs(player_url = robert_url, season_end_year = 2021, stat_type = "summary")
robert_stats_2022 <- fb_player_match_logs(player_url = robert_url, season_end_year = 2022, stat_type = "summary")
robert_stats_2023 <- fb_player_match_logs(player_url = robert_url, season_end_year = 2023, stat_type = "summary")
robert_combined_stats <- bind_rows(
  robert_stats_2021,
  robert_stats_2022,
  robert_stats_2023
)
head(robert_combined_stats)
```

### Analysis on Goal Expectancy (xG) vs Rest days before the game and Cumulative Minutes Played

```{r}

robert_combined_stats$Date <- as.Date(robert_combined_stats$Date, format = "%Y-%m-%d")

str(robert_combined_stats$Date) 

robert_combined_stats <- robert_combined_stats |>
  arrange(Date) |>
  mutate(
    Cumulative_Min = cumsum(Min),
    Rest_Days = as.numeric(difftime(Date, lag(Date, default = first(Date)), units = "days"))
  )
robert_combined_stats <- robert_combined_stats |>
  mutate(Rest_Days_Group = cut(Rest_Days, breaks = c(0, 3, 7, 14, Inf), labels = c("0-3 days", "4-7 days", "8-14 days", "15+ days")))
ggplot(robert_combined_stats, aes(x = Rest_Days_Group, y = xG_Expected)) +
  geom_violin(fill = "skyblue", alpha = 0.7) +
  geom_jitter(width = 0.1, color = "darkblue", alpha = 0.6) + 
  labs(
    title = "Distribution of Goal Expectancy (xG) by Rest Days Group",
    x = "Rest Days Group",
    y = "Goal Expectancy (xG)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  )
```

```{r}
robert_combined_stats <- robert_combined_stats |>
  arrange(Date) |>
  mutate(Cumulative_Min = cumsum(Min))

ggplot(robert_combined_stats, aes(x = Cumulative_Min, y = xG_Expected)) +
  geom_point(size = 3, color = "black") + 
  geom_smooth(method = "lm", se = FALSE, color = "blue", 
              linetype = "solid", size = 1, 
              aes(label = "Trend Line: Goal Expectancy")) +
  labs(
    title = "Cumulative Minutes vs. Goal Expectancy (xG)",
    x = "Cumulative Minutes",
    y = "Goal Expectancy (xG)"
  ) +
  theme_minimal() 
```

### Goal Expectancy (xG) on Cup Competition vs League Games:

```{r}
league_matches <- grep("Matchweek", robert_combined_stats$Comp, value = TRUE)

robert_combined_stats <- robert_combined_stats |>
  mutate(Competition_Type = case_when(
    Comp %in% league_matches ~ "League Match",
    Comp %in% c("UEFA Super Cup", "DFL-Supercup", "DFB-Pokal","WCQ", "Round of 16", "Quarter-finals", "UEFA Euro Qualifying",
                "Round of 32", "Semi-finals", "Final", "Knockout round play-offs", 
                "Second round", "First round", "UEFA Nations League", "Champions Lg") ~ "Cup Competition",
    Comp %in% c("League A", "Group stage", "Finals") ~ "International Competition",
    TRUE ~ "Other"
  ))

ggplot(robert_combined_stats, aes(x = xG_Expected, fill = Competition_Type)) +
  geom_density(alpha = 0.6) +
  labs(
    title = "Density Plot of Goal Expectancy (xG) by Competition Type",
    x = "Goal Expectancy (xG)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  )
```

### Shots on Target on Cup Competition vs League Games:

```{r}
ggplot(robert_combined_stats, aes(x = Competition_Type, y = SoT_Performance, fill = Competition_Type)) +
  geom_boxplot(alpha = 0.7, outlier.color = "red", outlier.size = 2) +  
  labs(
    title = "Shots on Target by Competition Type",
    x = "Competition Type",
    y = "Shots on Target"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
```

## Use Case 2: Injury Analysis of Spanish League Players from TransferMarkt

```{r}
injuries_df <- tm_league_injuries(country_name = "Spain")
head(injuries_df)
urls <- injuries_df$player_url
new_vars <- list()
for (i in seq_along(urls)) {
  new_vars[[paste0("df_", i)]] <- tm_player_injury_history(urls[i])
}
```

```{r}
combined_df <- data.frame()
for (i in 1:length(new_vars)) {
  combined_df <- rbind(combined_df, new_vars[[i]])
}
```

### Injury Types and Frequency 

```{r}
ggplot(combined_df, aes(x = injury)) +
  geom_bar() +
  labs(x = "Injury Type", y = "Count", title = "Injury Type Distribution") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
combined_df <- data.frame()

for (i in 1:length(new_vars)) {
  combined_df <- rbind(combined_df, new_vars[[i]])
}
combined_df$duration <- as.numeric(gsub(" days", "", combined_df$duration))
```

### Comparing Injury Duration Across Seasons

```{r}
ggplot(combined_df, aes(x = season_injured, y = duration)) +
  geom_boxplot() +
  labs(x = "Season", y = "Injury Duration", title = "Injury Duration by Season") +
  theme_minimal()
```

### Club-wise Injury Count

```{r}

club_injury_count <- combined_df |>
  group_by(club) |>
  summarise(injury_count = n())
club_injury_count <- club_injury_count |>
  arrange(desc(injury_count))
head(club_injury_count)
ggplot(club_injury_count, aes(x = reorder(club, -injury_count), y = injury_count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "Club", y = "Number of Injuries", title = "Injury Count by Club") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Injury Duration Patterns by Club

```{r fig.width=10, fig.height=8}
ggplot(combined_df, aes(x = reorder(club, -as.numeric(duration)), y = as.numeric(duration))) +
  geom_boxplot(fill = "lightgreen", color = "darkgreen") +
  labs(x = "Club", y = "Injury Duration (days)", title = "Distribution of Injury Duration by Club") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.margin = unit(c(1, 1, 1, 2), "cm") # Add extra space on the left
  )

```

### Club-wise Injury Trends Across Seasons

```{r}
injuries_by_club_season <- combined_df |>
  group_by(club, season_injured) |>
  summarise(injury_count = n(), .groups = "drop")

club_total_injury <- injuries_by_club_season |>
  group_by(club) |>
  summarise(total_injury = sum(injury_count))

injuries_by_club_season <- injuries_by_club_season |>
  group_by(club) |>
  mutate(total_injury_count = sum(injury_count)) |>
  ungroup()

fig <- plot_ly(injuries_by_club_season, 
               x = ~season_injured,  
               y = ~injury_count, 
               color = ~club,        
               type = 'bar', 
               hoverinfo = 'x+y+text',
               text = ~paste("Club: ", club, "<br>Injury Count: ", injury_count, "<br>Season: ", season_injured)) |>
  layout(title = "Injuries by Club and Season",
         xaxis = list(title = "Season", tickangle = 0),
         yaxis = list(title = "Number of Injuries"),
         legend = list(title = list(text = 'Club')),
         showlegend = TRUE)

fig <- fig |>
  layout(
    legend = list(
      itemclick = "toggleothers"  
    )
  )

fig
```

### Injury Analysis by Clubs over Time

```{r}
filtered_df <- combined_df |>
  group_by(club) |>
  filter(sum(!is.na(injured_since)) > 0) |>
  ungroup() |>
  filter(injured_since > 0)

p <- ggplot(filtered_df, aes(x = injured_since, color = club)) +
  geom_line(stat = "count", size = 1.5) +  
  geom_point(stat = "count", size = 3) +  
  labs(x = "Year", y = "Number of Injuries", title = "Trend of Injuries Over Time by Club") +
  theme_minimal() +
  theme(legend.position = "right", legend.text = element_text(size = 8))

plotly_plot <- ggplotly(p) |>
  layout(legend = list(itemclick = "toggleothers"))  

plotly_plot
```

## Use Case 3: Analysis on Teams in EPL (from fbref.com) 

```{r}
epl_team_possession <- fb_season_team_stats(
  country = "ENG", 
  gender = "M", 
  season_end_year = 2023, 
  tier = "1st", 
  stat_type = "possession"
)

unique(epl_team_possession$Squad)
```

```{r}
epl_team_possession$Squad <- as.character(epl_team_possession$Squad)

# Filter to keep rows where Squad does not contain "vs"
epl_team_possession <- epl_team_possession[!grepl("\\bvs\\b", epl_team_possession$Squad, ignore.case = TRUE), ]

epl_team_possession |>
  arrange(desc(Poss)) |>
  head(10) |>
  ggplot(aes(x = reorder(Squad, Poss), y = Poss)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() +
  labs(
    title = "Top 10 Teams by Possession in EPL 2021",
    x = "Team",
    y = "Possession (%)"
  ) +
  theme_minimal()
```

## Use Case 4: Team Defense analysis - Tackle success rate vs Total Tackles (from fbref.com) 

```{r}
epl_defense_stats <- fb_season_team_stats("ENG", "M", 2021, "1st", "defense")
epl_defense_stats <- epl_defense_stats[!grepl("\\bvs\\b", epl_defense_stats$Squad, ignore.case = TRUE), ]


ggplot(epl_defense_stats, aes(x = Tkl_Tackles, y = Tkl_percent_Challenges, size = Int, label = Squad)) +
  geom_point(color = "purple", fill = "lightblue", shape = 21, alpha = 0.6) +  
  geom_smooth(method = "lm", se = FALSE, color = "darkblue", linetype = "dashed") +  
  geom_text_repel(size = 3, max.overlaps = 15) + 
  scale_size_continuous(name = "Interceptions", range = c(3, 8)) +  
  labs(
    title = "Tackle Success Rate vs Total Tackles (Bubble Size = Interceptions)",
    x = "Total Tackles",
    y = "Tackle Success Rate (%)"
  ) +
  theme_minimal(base_size = 8) +  
  theme(
    plot.title = element_text(hjust = 0.5), 
    legend.position = "right"
  )
```
